Simple connectome inference from partial correlation statistics in calcium imaging

### 2014-07-28 response.tex

\textbf{A}: It is mentioned in the paper that there are enough samples to invert the covariance matrix. We have not done a statistical analysis in the scope of this work because we only use the partial correlation to rank pairs and therefore a statistical analysis is not really needed to make that ranking. However, it would be interesting to do it in future works.

\section*{Reviewer 2:}

\textcolor{blue}{It would be necessary to have information about the computational complexity, scalability, memory requirements, and execution time for these methods.}

\textbf{A}: We obviously agree with the reviewer that such informations are interesting, especially in the context of a challenge. For the simplified method, it only takes a few hours (about 3 to 6 hours depending on the machine used) and a few gigabytes of memory. The ``Full method'' needs the double of time (about 6 to 12 hours) and it requires a few hours more to take into account the edge directionality. GTE and GENIE3 take more time to compute than PC. However, GENIE3 is highy parallelizable (by taking several smaller forests or applying GENIE3 on variable subsets) and the execution time can be significantly reduced.

@@

However, let us note that the simple method presented in the paper already gives good performances despite its light computational needs.

We also would like to mention that our code is available publicly on \url{https://github.com/asutera/kaggle-connectomics} and that some verifications, including memory requirements and execution time, of the ``Full method'' are displayed on the challenge website \url{http://connectomics.chalearn.org/home/challenge-results-verification}.\\

% \textcolor{blue}{- Soundness (good experiments, good proofs):  4\\}
\noindent
\textcolor{blue}{In section 4, PC, GTE, and GENIE3 are compared with other methods. Another worthwhile experiment would be to extract features using the procedure described by the authors, do a Principal Component Analysis(PCA), and then apply PC, etc. other methods.}

\textbf{A}: We are not sure what exactly suggests the reviewer but applying PCA before PC, GTE and GENIE3 is not straightforward. Let us note that for each method we choose the preprocessing which yields the best performance.\\

### 2014-07-28 article.tex
6 pages constraint + caps
last changes

\usepackage{enumerate}
\usepackage{bbold}
\usepackage{booktabs}
\usepackage{natbib}
\usepackage[utf8]{inputenc}
\usepackage[super]{nth}

@@

\vspace{-0.8cm}
\begin{abstract}
In this work, we propose a simple yet effective solution to the problem of
connectome inference in calcium imaging data. The proposed algorithm consists of
two steps. First, processing the raw signals to detect neural peak activities.
Second, inferring the degree of association between neurons from partial
correlation statistics. This paper summarises the methodology that led us to
win the Connectomics Challenge, proposes a simplified version of our method, and
finally compares our results with respect to other inference methods.
\end{abstract}

@@

\section{Experiments} \label{sec:results}

\paragraph{Data and evaluation metrics.}

@@

\caption{Top: Performance on \textit{normal-1,2,3,4} with partial correlation and different filtering functions.
Bottom: Performance on \textit{normal-1,2,3,4} with different methods.}
\label{tab:comparison}
\centering
\small
\begin{tabular}{| l | c c c c | c c c c |}
\hline
& \multicolumn{4}{c|}{AUROC} & \multicolumn{4}{c|}{AUPRC} \\
\textit{Method} $\backslash$ \textit{normal-} & \textit{1} & \textit{2} & \textit{3} & \textit{4} & \textit{1} & \textit{2} & \textit{3} & \textit{4} \\
\hline
\hline
No  filtering       					& 0.777 & 0.767 & 0.772 & 0.774 & 0.070 & 0.064 & 0.068 & 0.072\\
$ h \circ g \circ f_1$                  & 0.923 & 0.925 & 0.923 & 0.922 & 0.311 & 0.315 & 0.313 & 0.304\\
$ w \circ h \circ g \circ f_1$          & 0.931 & 0.929 & 0.928 & 0.926 & 0.326 & 0.323 & 0.319 & 0.303\\
+ PCA         							& 0.932 & 0.930 & 0.928 & 0.926 & 0.355 & 0.353 & 0.350 & 0.333\\
Averaging           					& 0.937 & 0.935 & 0.935 & 0.931 & 0.391 &  0.390 &  0.385 & 0.375\\
Full method           					& \textbf{0.943} & \textbf{0.942} & \textbf{0.942} & \textbf{0.939} & \textbf{0.403} & \textbf{0.404} & \textbf{0.398} & \textbf{0.388}\\
\hline
PC & 0.886 & 0.884 & 0.891 &  0.877 & 0.153 & 0.145 & 0.170 & 0.132\\
GTE & 0.890 & 0.893 & 0.894 & 0.873 & 0.171 & 0.174 & 0.197 & 0.142\\
GENIE3 & 0.892 & 0.891 & 0.887 & 0.887 & 0.232 & 0.221 & 0.237 & 0.215 \\
\hline
\end{tabular}

\end{table}

@@

The fact that our method is unable to predict edge directions does not seem to
be a disadvantage with respect to GTE and GENIE3. Although partial correlation
scores each edge, and its opposite, similarly, it can reach precision values
higher than 0.5 (see Figure \ref{fig:curves}(b)), suggesting that it mainly ranks high
pairs of neurons that interact in both directions.  It is interesting also to
note that, on \textit{normal-2}, a method that perfectly predicts the
undirected network (i.e., that gives a score of $1$ to each pair $(i,j)$ such that
$(i,j)\in E$ or $(j,i)\in E$, and $0$ otherwise) already reaches an AUROC as high
as $0.995$ and an AUPRC of $0.789$.

\begin{figure}[t]
\centering
\subfigure[ROC curves]{\includegraphics[width=0.45\textwidth]{curve_roc} \label{fig:roc_curve}}
\subfigure[Precision-recall curves]{\includegraphics[width=0.45\textwidth]{curve_pr} \label{fig:pr_curve}}
\caption{ROC (left) and PR (right) curves on \textit{normal-2} for the compared methods. Areas under the curves are reported in the legend.}
\label{fig:curves}
\end{figure}

@@

\begin{table}[h]
 \caption{Connectomics challenge summary.}
 \label{tab:summary}
 \centering
 \small
 \begin{tabular}{| l|c|} \hline
 Team Name & The AAAGV Team \\ \hline
 Private leaderboard position & \nth{1} \\ \hline
 Private leaderboard performance & 0.94161 \\ \hline
 Private leaderbard performance of the winner & / \\ \hline
 \end{tabular}
 \end{table}

>>>>

\usepackage{enumerate}
\usepackage{bbold}
\usepackage{booktabs}
\usepackage{natbib}
\usepackage[utf8]{inputenc}
\usepackage[super]{nth}
\setlength{\textfloatsep}{15pt}

@@

\vspace{-1cm}
\begin{abstract}
In this work, we propose a simple yet effective solution to the problem of
connectome inference in calcium imaging data. The proposed algorithm consists of
two steps. First, processing the raw signals to detect neural peak activities.
Second, inferring the degree of association between neurons from partial
correlation statistics. This paper summarises the methodology that led us to
win the Connectomics Challenge, proposes a simplified version of our method, and
finally compares our results with respect to other inference methods.
\end{abstract}

@@

\section{Experiments} \label{sec:results}
\paragraph{Data and evaluation metrics.}

@@

\caption{Top: Performance on \textit{normal-1,2,3,4} with partial correlation and different filtering functions.
Bottom: Performance on \textit{normal-1,2,3,4} with different methods.}
\label{tab:comparison}
\centering
\small
\begin{tabular}{| l | c c c c | c c c c |}
\hline
& \multicolumn{4}{c|}{AUROC} & \multicolumn{4}{c|}{AUPRC} \\
\textit{Method} $\backslash$ \textit{normal-} & \textit{1} & \textit{2} & \textit{3} & \textit{4} & \textit{1} & \textit{2} & \textit{3} & \textit{4} \\
\hline
\hline
No  filtering       					& 0.777 & 0.767 & 0.772 & 0.774 & 0.070 & 0.064 & 0.068 & 0.072\\
$ h \circ g \circ f_1$                  & 0.923 & 0.925 & 0.923 & 0.922 & 0.311 & 0.315 & 0.313 & 0.304\\
$ w \circ h \circ g \circ f_1$          & 0.931 & 0.929 & 0.928 & 0.926 & 0.326 & 0.323 & 0.319 & 0.303\\
+ PCA         							& 0.932 & 0.930 & 0.928 & 0.926 & 0.355 & 0.353 & 0.350 & 0.333\\
Averaging           					& 0.937 & 0.935 & 0.935 & 0.931 & 0.391 &  0.390 &  0.385 & 0.375\\
Full method           					& \textbf{0.943} & \textbf{0.942} & \textbf{0.942} & \textbf{0.939} & \textbf{0.403} & \textbf{0.404} & \textbf{0.398} & \textbf{0.388}\\
\hline
PC & 0.886 & 0.884 & 0.891 &  0.877 & 0.153 & 0.145 & 0.170 & 0.132\\
GTE & 0.890 & 0.893 & 0.894 & 0.873 & 0.171 & 0.174 & 0.197 & 0.142\\
GENIE3 & 0.892 & 0.891 & 0.887 & 0.887 & 0.232 & 0.221 & 0.237 & 0.215 \\
\hline
\end{tabular}
\end{table}

@@

The fact that our method is unable to predict edge directions does not seem to
be a disadvantage with respect to GTE and GENIE3. Although partial correlation
scores each edge, and its opposite, similarly, it can reach precision values
higher than 0.5 (see Figure \ref{fig:curves}(b)), suggesting that it mainly ranks high
pairs of neurons that interact in both directions.  It is interesting also to
note that, on \textit{normal-2}, a method that perfectly predicts the
undirected network (i.e., that gives a score of $1$ to each pair $(i,j)$ such that
$(i,j)\in E$ or $(j,i)\in E$, and $0$ otherwise) already reaches an AUROC as high
as $0.995$ and an AUPRC of $0.789$.
\begin{figure}[t]
\centering
\subfigure[ROC curves]{\includegraphics[width=0.45\textwidth]{curve_roc} \label{fig:roc_curve}}
\subfigure[Precision-recall curves]{\includegraphics[width=0.45\textwidth]{curve_pr} \label{fig:pr_curve}}
\caption{ROC (left) and PR (right) curves on \textit{normal-2} for the compared methods. Areas under the curves are reported in the legend.}
\label{fig:curves}
\end{figure}

@@

\begin{table}[h]
 \caption{Connectomics Challenge summary.}
 \label{tab:summary}
 \centering
 \small
 \begin{tabular}{| l|c|} \hline
 Team Name & The AAAGV Team \\ \hline
 Private leaderboard position & \nth{1} \\ \hline
 Private leaderboard performance & 0.94161 \\ \hline
 Private leaderboard performance of the winner & idem \\ \hline
 \end{tabular}
 \end{table}
